---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

For this assignment, I need the packages tidyverse and lubridate. At first, the packages are loaded, then the activity file is unzipped and assigned to variables. The date-column needs to be formatted into a date variable.

```{r echo = TRUE}
library(tidyverse)
library(lubridate)


unzip("./activity.zip")
activityfile <- read.csv("activity.csv")
activityfile$date <- as.Date(activityfile$date)
```

## What is mean total number of steps taken per day?


```{r echo = TRUE}
averagesteps <- mean(activityfile$steps, na.rm = TRUE)
mediansteps <- median(activityfile$steps)
```

The mean total number of steps taken per day is `r averagesteps`
The median number of steps taken per day is `r mediansteps`

## What is the average daily activity pattern?

```{r echo = TRUE}
stepsperdaymean <- aggregate(activityfile$steps, list(activityfile$date), FUN = mean, na.rm = TRUE)
names(stepsperdaymean) <- c("Date", "Steps.per.day.mean")
stepsperdaymean$Date <- as.Date(stepsperdaymean$Date)
stepsperdaymean

ggplot(data = stepsperdaymean, aes(x = Date, y = Steps.per.day.mean, group = 1)) +
  geom_point() + geom_line()

```

#Q4:The 5-minute interval that, on average, contains the maximum number of steps
```{r echo = TRUE}
activityfileomitted <- na.omit(activityfile)
fiveinterval <- activityfileomitted$interval[activityfileomitted$steps == as.integer(max(activityfileomitted$steps))]
```

The 5-minute interval on average containing the maximum number of steps is `r fiveinterval`


## Imputing missing values
```{r echo = TRUE}
nafiles <- sum(is.na(activityfile$steps))
activityfile$steps[is.na(activityfile$steps)] <- mean(activityfile$steps[activityfile$date])
```
The numer of NA's is `r nafiles`
The missing files are filled with the mean values.


## Are there differences in activity patterns between weekdays and weekends?
```{r echo = TRUE}
activityfile$weekday <- weekdays(activityfile$date)
activityfile$weekend <- "weekend"
activityfile$weekend[activityfile$weekday == "Montag" | activityfile$weekday == "Dienstag" | activityfile$weekday == "Mittwoch" | activityfile$weekday == "Donnerstag" | activityfile$weekday == "Freitag"] <- "weekdays"
stepsperintervalmean <- aggregate(activityfile$steps ~ activityfile$interval+activityfile$weekend, data = activityfile,FUN =  mean)
names(stepsperintervalmean) <- c("Interval", "Weekpart", "Steps")
ggplot(stepsperintervalmean, aes(Interval, Steps)) + geom_point() +
  facet_grid(rows = vars(Weekpart))
```

There are some differences, e.g. higher number of steps per day in the morning during weekdays, but a higher number during the afternoon during the weekend.
